---
title: "General Internal Medicine Shared Admissions Workload 2022"
author:
  - name: Dr Saiful Safuan Md Sani 
    url: https://saifulsafuan.com
    affiliation: Clinical Research Centre & Department of Medicine Hospital Kuala Lumpur
  - name: Dr Yoon Chee Kin
    affiliation: Clinical Research Centre & Department of Medicine Hospital Pulau Pinang
  - name: Dr Cheah Wee Kooi
    affiliation: Clinical Research Centre & Department of Medicine Hospital Taiping
date: "`2023-09-12`"
output:
  distill::distill_article:
    self_contained: yes
    toc: yes
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
# first pub 17 Jan 2022
library(knitr)
library(distill)
library(googlesheets4)
library(gsheet)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
library(RColorBrewer)
#library(kableExtra)
#library(zoo) # moving averages
#library(RcppRoll) # running total

#  distill::distill_article:
#    self_contained: yes
#    toc: yes
#    toc_depth: 2
#    _includes: analytics.html

```

```{r url shared explorer adm, include=FALSE}
url <- "https://docs.google.com/spreadsheets/d/1z_1ydaDB0Fhz5FJ2RUtfKb_0bYafuhtebDP3y9L2-vo/edit#gid=1570948431"  

explorer_shared <- gsheet2tbl(url, "explorer_shared")

explorer_shared$type <- as.factor(explorer_shared$type)
explorer_shared$type <- relevel(explorer_shared$type, "STATE")

explorer_shared$year <- as.factor(explorer_shared$year)
explorer_shared$year <- relevel(explorer_shared$year, "2019")

glimpse(explorer_shared)

col3 <- c(`STATE` = "#C60000", 
          `MAJOR` = "#1E434C", 
          `MINOR` = "#7AA802", 
          `NONSPECIALIST` = "#C99E10")

#dataexplorer_shared <- arrange(explorer_shared, desc(spec_adequacy))
```
# General Medical Admissions 2019 (Shared Admissions Workload)

```{r adm2019, fig.cap="General Medical Admissions 2019 (Shared Admissions Workload).", out.extra="class=external",  fig.width=10, fig.height=10, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(year == '2019')

figs <- fig %>% plot_ly(x = ~adm, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "General Medical Admissions 2019",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 8),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

# Number of Specialists 2019
```{r specnshare2019, fig.cap="Number of Specialists 2019 (Shared Admissions Workload).", out.extra="class=external",  fig.width=10, fig.height=10, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(year == '2019')
fig <- arrange(fig, desc(spec_n_share))

figs <- fig %>% plot_ly(x = ~spec_n_share, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Number of Specialists 2019",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 8),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

# Specialists per 1000 Admissions 2019

```{r specper1000adm2019, fig.cap="Specialist per 1000 Admissions 2019 (Shared Admissions Workload).", out.extra="class=external",  fig.width=10, fig.height=10, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(year == '2019')
fig <- arrange(fig, desc(spec_nshare_per_1000_adm))

figs <- fig %>% plot_ly(x = ~spec_nshare_per_1000_adm, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Specialist per 1000 Admissions 2019",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 8),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

# Specialists per 1000 Admissions 2019 by Type of Facility

```{r specper1000adm2019facility, fig.cap="Specialist per 1000 Admissions 2019 (Shared Admissions Workload).", out.extra="class=external",  fig.width=10, fig.height=10, message=FALSE, echo=FALSE, warning=FALSE}

states <- explorer_shared %>% filter(type == 'STATE', year == '2019')
states$hospital <- factor(states$hospital)
#str(states)

p_states <- plot_ly(states, y = ~hospital, x = ~spec_nshare_per_1000_adm , type = 'bar',  name = 'States', color = ~factor(type), colors = col3)
p_states <- p_states %>% add_trace(states, y = ~hospital, x = ~median, 
                      type = 'scatter', 
                      name = 'Median',
                      mode = "lines", 
                      line = list(color = "crimson", dash = 'dot', width = 7), 
                      opacity = 0.97)
p_states <- p_states %>% layout(xaxis = list(title = 'Number of specialists per 1000 admissions'), yaxis = list(title = '', categoryorder = "total ascending", tickfont = list(size = 8)))

#p_states

major <- explorer_shared %>% filter(type == 'MAJOR', year == '2019')
major$hospital <- factor(major$hospital)
#str(major)

p_major <- plot_ly(major, y = ~hospital, x = ~spec_nshare_per_1000_adm, type = 'bar',  name = 'Major', color = ~factor(type), colors = col3)
p_major <- p_major %>% add_trace(major, y = ~hospital, x = ~median, 
                      type = 'scatter', 
                      name = 'Median',
                      mode = "lines", 
                      line = list(color = "crimson", dash = 'dot', width = 7), 
                      opacity = 0.97)
p_major <- p_major %>% layout(xaxis = list(title = 'Number of specialists per 1000 admissions'), yaxis = list(title = '', categoryorder = "total ascending", tickfont = list(size = 8)))

#p_major

minor <- explorer_shared %>% filter(type == 'MINOR', year == '2019')
minor$hospital <- factor(minor$hospital)
#str(minor)

p_minor <- plot_ly(minor, y = ~hospital, x = ~spec_nshare_per_1000_adm, type = 'bar',  name = 'Minor', color = ~factor(type), colors = col3)
p_minor <- p_minor %>% add_trace(minor, y = ~hospital, x = ~median, 
                      type = 'scatter', 
                      name = 'Median',
                      mode = "lines", 
                      line = list(color = "crimson", dash = 'dot', width = 7), 
                      opacity = 0.97)
p_minor <- p_minor %>% layout(xaxis = list(title = 'Number of specialists per 1000 admissions'), yaxis = list(title = '', categoryorder = "total ascending", tickfont = list(size = 8)))

#p_minor
```

```{r bytype2, fig.cap="Specialist per 1000 Admissions 2019 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}
subplot(p_states, p_major, p_minor , nrows = 3, heights = c(0.25, 0.5, 0.25), shareX = TRUE, shareY = TRUE) %>% 
  layout(title = list(text = "Specialist per 1000 Admissions 2019 (Shared Admissions Workload)."))

#options(browser = 'false')
#api_create(p, filename = "hosp2019_spec_n_per_1000_adm")
```
<br>

<a href="#top">Return to table of contents</a>

# Specialist Adequacy Calculation

<br> 

\begin{equation}
{Adequacy = {Number\;of\;specialists}_{Actual} - {Number\;of\;specialists}_{Median}}
(\#eq:Adeq)
\end{equation}

<br>

derived from 

\begin{equation}
{{Number\;of\; specialists}_{Median} = {Manpower\;Rate}_{Median} \times  \left( \displaystyle \frac{Admissions}{1000}\right) }
(\#eq:MRn)
\end{equation} 

<br>

derived from

\begin{equation}
{Manpower\;Rate = {Number\;of\; specialists} \div \left( \displaystyle \frac{Admissions}{1000}\right) }
(\#eq:MR)
\end{equation} 

<br>

<a href="#top">Return to table of contents</a>

\pagebreak

# Specialist Excess / Shortages 2019

```{r adequacy2019, fig.cap="Specialist Excess / Shortages 2019 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(year == '2019')
fig <- arrange(fig, desc(spec_adequacy))

figs <- fig %>% plot_ly(x = ~spec_adequacy, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Specialist Excess / Shortages 2019",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

# Specialist Excess / Shortages 2019:2022
## State

```{r xsstate, fig.cap="Specialist Excesses and Shortages at State Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'STATE')
fig <- arrange(fig, year, desc(spec_adequacy))

figs <- fig %>% plot_ly(x = ~spec_adequacy, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Specialist Excesses and Shortages at State Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

## Major

```{r xsmajor, fig.cap="Specialist Excesses and Shortages at Major Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'MAJOR')
fig <- arrange(fig, year, desc(spec_adequacy))

figs <- fig %>% plot_ly(x = ~spec_adequacy, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Specialist Excesses and Shortages at Major Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

## Minor

```{r xsminor, fig.cap="Specialist Excesses and Shortages at Minor Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'MINOR')
fig <- arrange(fig, year, desc(spec_adequacy))

figs <- fig %>% plot_ly(x = ~spec_adequacy, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Specialist Excesses and Shortages at Minor Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

# Number of Specialists 2019:2022
## State

```{r nstate, fig.cap="Number of Specialists at State Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external", fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'STATE')
fig <- arrange(fig, year, desc(spec_n_share))

figs <- fig %>% plot_ly(x = ~spec_n_share, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Number of Specialists at State Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

## Major

```{r nmajor, fig.cap="Number of Specialists at Major Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external", fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'MAJOR')
fig <- arrange(fig, year, desc(spec_n_share))

figs <- fig %>% plot_ly(x = ~spec_n_share, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Number of Specialists at Major Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

## Minor

```{r nminor, fig.cap="Number of Specialists at Minor Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external", fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'MINOR')
fig <- arrange(fig, year, desc(spec_n_share))

figs <- fig %>% plot_ly(x = ~spec_n_share, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "Number of Specialists at Minor Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

# General Medical Admissions 2019:2022
## State

```{r admstate, fig.cap="General Medical Admissions at State Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'STATE')
fig <- arrange(fig, year, desc(adm))

figs <- fig %>% plot_ly(x = ~adm, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "General Medical Admissions at State Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

## Major

```{r admmajor, fig.cap="General Medical Admissions at Major Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'MAJOR')
fig <- arrange(fig, year, desc(adm))

figs <- fig %>% plot_ly(x = ~adm, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "General Medical Admissions at Major Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>

## Minor

```{r admminor, fig.cap="General Medical Admissions at Minor Hospitals 2019 to 2022 (Shared Admissions Workload).", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

fig <- explorer_shared
fig <- fig %>% filter(type == 'MINOR')
fig <- arrange(fig, year, desc(adm))

figs <- fig %>% plot_ly(x = ~adm, y = ~hospital, color = ~type, colors = col3, symbol = ~year, type = "scatter", mode = "markers", showlegend = T) %>% layout(
     title = "General Medical Admissions at Minor Hospitals 2019 to 2022",
     yaxis = list(title = "",
     categoryarray = ~fig$hospital, tickfont = list(size = 12),
     categoryorder = "array",
     autorange = "reversed"),
     xaxis = list(title = "Number of specialists"),
     legend = list(x = 0.97, y = 0.95)
     )

figs
```
<br>

<a href="#top">Return to table of contents</a>