---
title: "General Internal Medicine Shared Admissions Workload 2022"
author:
  - name: Dr Saiful Safuan Md Sani 
    url: https://saifulsafuan.com
    affiliation: Clinical Research Centre & Department of Medicine Hospital Kuala Lumpur
  - name: Dr Yoon Chee Kin
    affiliation: Clinical Research Centre & Department of Medicine Hospital Pulau Pinang
date: "`2023-09-01`"
output:
  distill::distill_article:
    self_contained: yes
    toc: yes
    toc_float: yes
always_allow_html: yes
---


```{r setup, include=FALSE}
# first pub 17 Jan 2022
library(knitr)
library(distill)
library(googlesheets4)
library(gsheet)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
#library(kableExtra)
#library(zoo) # moving averages
#library(RcppRoll) # running total

#  distill::distill_article:
#    self_contained: yes
#    toc: yes
#    toc_depth: 2
#    _includes: analytics.html

```
```{r urlshare, include=FALSE}
urlshare <- "https://docs.google.com/spreadsheets/d/1z_1ydaDB0Fhz5FJ2RUtfKb_0bYafuhtebDP3y9L2-vo/edit#gid=1618837026" ### 
hosp2022share <- gsheet2tbl(urlshare, "adm") 
glimpse(hosp2022share)
```

```{r include=FALSE}
hosp2022share$type <- as.factor(hosp2022share$type)
hosp2022share$type <- relevel(hosp2022share$type, "STATE")

rownames(hosp2022share) <- hosp2022share$hospital
rownames(hosp2022share)

#col3 <- c(`state` = "#150377", `major` = "salmon", `minor` = "grey")

hosp2022share$hospital <- factor(hosp2022share$hospital,
                   levels = hosp2022share$hospital)

datashare <- arrange(hosp2022share, type, desc(spec_nshare_per_1000_adm))
```
## Shared General Medical Admissions Workload
```{r, fig.cap="General Medical Admissions.", out.extra="class=external",  fig.width=8, fig.height=10, message=FALSE, echo=FALSE, warning=FALSE}
#### Current workload
col3 <- c(`STATE` = "#150377", `MAJOR` = "salmon", `MINOR` = "grey", `NONSPECIALIST` = "light blue")

datashare_adm <- arrange(hosp2022share, adm)
datashare_adm$hospital <- factor(datashare_adm$hospital,
                   levels = datashare_adm$hospital)

p <- plot_ly(datashare_adm, y = ~factor(hospital), x = ~adm, alpha = 0.97, size = 7)
subplot(
  add_bars(p, color = ~factor(type), colors = col3,  showlegend = TRUE) 
) %>% layout(
     title = "Shared General Medical Admissions Workload<br><sup></sup>",
     yaxis = list(title = "",
     categoryorder = "array",
     categoryarray = ~datashare_adm$hospital, tickfont = list(size = 10)),
     xaxis = list(title = "Number of admissions")
) %>% layout(legend = list(x = 0.77, y = 0.9)) %>% 
 layout(annotations = 
 list(x = 1, y = 0.51, text = "", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(size=12, color="grey"))
 )


#options(browser = 'false')
#api_create(p, filename = "hosp2019_admission")
```
<a href="#top">Return to table of contents</a>

\pagebreak

## Specialist (per 1000 Admissions) Handling Shared General Medical Admissions Workload by Type of Facility 2022
```{r, fig.cap="Specialist per 1000 Admissions.", out.extra="class=external", message=FALSE, echo=FALSE, warning=FALSE}
states <- datashare %>% filter(type == 'STATE')
states$hospital <- factor(states$hospital)
#str(states)

p_states <- plot_ly(states, y = ~hospital, x = ~spec_nshare_per_1000_adm, type = 'bar',  name = 'States', color = ~factor(type), colors = col3)
p_states <- p_states %>% add_trace(states, y = ~hospital, x = ~median, 
                      type = 'scatter', 
                      name = 'Median',
                      mode = "lines", 
                      line = list(color = "crimson", dash = 'dot', width = 7), 
                      opacity = 0.97)
p_states <- p_states %>% layout(xaxis = list(title = 'Number of specialists per 1000 admissions'), yaxis = list(title = '', categoryorder = "total ascending", tickfont = list(size = 8)))

#p_states

major <- datashare %>% filter(type == 'MAJOR')
major$hospital <- factor(major$hospital)
#str(major)

p_major <- plot_ly(major, y = ~hospital, x = ~spec_nshare_per_1000_adm, type = 'bar',  name = 'Major', color = ~factor(type), colors = col3)
p_major <- p_major %>% add_trace(major, y = ~hospital, x = ~median, 
                      type = 'scatter', 
                      name = 'Median',
                      mode = "lines", 
                      line = list(color = "crimson", dash = 'dot', width = 7), 
                      opacity = 0.97)
p_major <- p_major %>% layout(xaxis = list(title = 'Number of specialists per 1000 admissions'), yaxis = list(title = '', categoryorder = "total ascending", tickfont = list(size = 8)))

#p_major

minor <- datashare %>% filter(type == 'MINOR')
minor$hospital <- factor(minor$hospital)
#str(minor)

p_minor <- plot_ly(minor, y = ~hospital, x = ~spec_nshare_per_1000_adm, type = 'bar',  name = 'Minor', color = ~factor(type), colors = col3)
p_minor <- p_minor %>% add_trace(minor, y = ~hospital, x = ~median, 
                      type = 'scatter', 
                      name = 'Median',
                      mode = "lines", 
                      line = list(color = "crimson", dash = 'dot', width = 7), 
                      opacity = 0.97)
p_minor <- p_minor %>% layout(xaxis = list(title = 'Number of specialists per 1000 admissions'), yaxis = list(title = '', categoryorder = "total ascending", tickfont = list(size = 8)))

#p_minor

nonspecialist <- datashare %>% filter(type == 'NONSPECIALIST')
nonspecialist$hospital <- factor(nonspecialist$hospital)
#str(nonspecialist)

p_nonspecialist <- plot_ly(nonspecialist, y = ~hospital, x = ~spec_nshare_per_1000_adm, type = 'bar',  name = 'Nonspecialist', color = ~factor(type), colors = col3)
p_nonspecialist <- p_nonspecialist %>% add_trace(nonspecialist, y = ~hospital, x = ~median, 
                      type = 'scatter', 
                      name = 'Median',
                      mode = "lines", 
                      line = list(color = "crimson", dash = 'dot', width = 7), 
                      opacity = 0.97)
p_nonspecialist <- p_nonspecialist %>% layout(xaxis = list(title = 'Number of specialists per 1000 admissions'), yaxis = list(title = '', categoryorder = "total ascending", tickfont = list(size = 8)))

#p_nonspecialist
```

```{r, fig.cap="Specialist (per 1000 Admissions) Handling Shared General Medical Admissions Workload 2022.", out.extra="class=external",  fig.width=8, fig.height=10, message=FALSE, echo=FALSE, warning=FALSE}
subplot(p_states, p_major, p_minor, nrows = 3, heights = c(0.25, 0.5, 0.25), shareX = TRUE, shareY = TRUE) %>% 
  layout(title = list(text = "Specialist (per 1000 Admissions) by Type of Facility 2022"))

#options(browser = 'false')
#api_create(p, filename = "hosp2019_spec_n_per_1000_adm")
```

## Specialist Adequacy Calculation

<br> 

\begin{equation}
{Adequacy = {Number\;of\;specialists}_{Actual} - {Number\;of\;specialists}_{Median}}
(\#eq:Adeq)
\end{equation}

<br>

derived from 

\begin{equation}
{{Number\;of\; specialists}_{Median} = {Manpower\;Rate}_{Median} \times  \left( \displaystyle \frac{Admissions}{1000}\right) }
(\#eq:MRn)
\end{equation} 

<br>

derived from

\begin{equation}
{Manpower\;Rate = {Number\;of\; specialists} \div \left( \displaystyle \frac{Admissions}{1000}\right) }
(\#eq:MR)
\end{equation} 

<br>

<a href="#top">Return to table of contents</a>

\pagebreak

## Specialist Adequacy 2022
```{r, fig.cap="\\label{fig:myfig4}Specialist Excesses and Shortages. Negative values are shortages, and positive values are excesses. See equations 1-3 for formulae to derive values.", out.extra="class=external",  fig.width=8, fig.height=8, message=FALSE, echo=FALSE, warning=FALSE}

col3 <- c(`STATE` = "#150377", `MAJOR` = "salmon", `MINOR` = "grey", `NONSPECIALIST` = "light blue")

adeq <- arrange(hosp2022share, spec_adequacy)
adeq$hospital <- factor(adeq$hospital,
                   levels = adeq$hospital)

p <- plot_ly(adeq, y = ~factor(hospital), x = ~spec_adequacy)
subplot(
  add_markers(p, color = ~factor(type), colors = col3,  showlegend = TRUE, alpha = 0.97, size = 13) 
) %>% layout(
     title = "Specialist Excesses and Shortages 2022",
     yaxis = list(title = "",
     categoryorder = "array",
     categoryarray = ~adeq$hospital, tickfont = list(size = 8)),
     xaxis = list(title = "Number of specialists")
) %>% layout(legend = list(x = 0.77, y = 0.8))

```
<br>

<a href="#top">Return to table of contents</a>